---
title: "Real Data Applications: Norwegian Infant Gut Microbiome Data"
author: 
  - Huang Lin$^1$
  - $^1$NICHD, 6710B Rockledge Dr, Bethesda, MD 20892
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document: 
    toc: true
    theme: united
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, 
                      message = FALSE, comment = NA,
                      fig.width = 6.25, fig.height = 5)

library(tidyverse)
library(microbiome)
library(randomcoloR)
library(ggpubr)
library(SpiecEasi)
library(openxlsx)
# library(energy)
library(robustbase)
library(DT)
options(DT.options = list(
  initComplete = JS("function(settings, json) {",
  "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});","}")))

source("../programs/00_secom.R")
```

```{r helper}
get_upper_tri = function(cormat){
    cormat[lower.tri(cormat)] = NA
    diag(cormat) = NA
    return(cormat)
}

hard_thresh = function(R, th){
  R_th = R
  R_th[abs(R) <= th] = 0
  return(R_th)
}
```

# 1. Data import {.tabset}

## Generate phyloseq objects

```{r, eval=FALSE}
# The Nomic data are not publicly available

# Metadata
meta_data = read_csv("../data/nomic/nomic_meta.csv")

# Taxonomy
tax = read_tsv("../data/nomic/nomic_taxnomy.txt") %>% 
  arrange(otu_id)
otu_id = tax$otu_id
tax = data.frame(tax[, -1]) %>%
  rowwise() %>%
  dplyr::mutate_all(function(x) strsplit(x, "__")[[1]][2]) %>%
  mutate(
    species = ifelse(!is.na(species) & !is.na(genus),
                     paste(strsplit(genus, "")[[1]][1], species, sep = "."),
                     NA)) %>%
  ungroup()
tax = as.matrix(tax)
rownames(tax) = otu_id

# Day 30

# OTU table
otu_table = read_tsv("../data/nomic/nomic_otu_30.txt")
sample_id = otu_table$`#SampleID`
otu_id = colnames(otu_table)[-1]
otu_table = t(otu_table[, -1])
otu_table = data.frame(otu_table)
rownames(otu_table) = otu_id
colnames(otu_table) = sample_id
otu_table = as.matrix(otu_table)

# OTU object
OTU = otu_table(otu_table, taxa_are_rows = TRUE)
META = sample_data(meta_data)
sample_names(META) = meta_data$`#SampleID`
TAX = tax_table(tax)
otu_data30 = phyloseq(OTU, TAX, META)

# Aggregate to phylum level
phylum_data30 = aggregate_taxa(otu_data30, "phylum")
phylum_data30 = subset_taxa(phylum_data30, phylum != "Unknown")

# Aggregate to family level
family_data30 = aggregate_taxa(otu_data30, "family")
family_data30 = subset_taxa(family_data30, family != "Unknown")

# Day 120 

# OTU table
otu_table = read_tsv("../data/nomic/nomic_otu_120.txt") 
sample_id = otu_table$`#SampleID`
otu_id = colnames(otu_table)[-1]
otu_table = t(otu_table[, -1])
otu_table = data.frame(otu_table)
rownames(otu_table) = otu_id
colnames(otu_table) = sample_id
otu_table = as.matrix(otu_table)

# OTU object
OTU = otu_table(otu_table, taxa_are_rows = TRUE)
META = sample_data(meta_data)
sample_names(META) = meta_data$`#SampleID`
TAX = tax_table(tax)
otu_data120 = phyloseq(OTU, TAX, META)

# Aggregate to phylum level
phylum_data120 = aggregate_taxa(otu_data120, "phylum")
phylum_data120 = subset_taxa(phylum_data120, phylum != "Unknown")

# Aggregate to family level
family_data120 = aggregate_taxa(otu_data120, "family")
family_data120 = subset_taxa(family_data120, family != "Unknown")

# Day 365 

# OTU table
otu_table = read_tsv("../data/nomic/nomic_otu_365.txt") 
sample_id = otu_table$`#SampleID`
otu_id = colnames(otu_table)[-1]
otu_table = t(otu_table[, -1])
otu_table = data.frame(otu_table)
rownames(otu_table) = otu_id
colnames(otu_table) = sample_id
otu_table = as.matrix(otu_table)

# OTU object
OTU = otu_table(otu_table, taxa_are_rows = TRUE)
META = sample_data(meta_data)
sample_names(META) = meta_data$`#SampleID`
TAX = tax_table(tax)
otu_data365 = phyloseq(OTU, TAX, META)

# Aggregate to phylum level
phylum_data365 = aggregate_taxa(otu_data365, "phylum")
phylum_data365 = subset_taxa(phylum_data365, phylum != "Unknown")

# Aggregate to family level
family_data365 = aggregate_taxa(otu_data365, "family")
family_data365 = subset_taxa(family_data365, family != "Unknown")
```

## Quality check

```{r, eval=FALSE}
# Sample size
df_sample = data.frame(day30 = nsamples(otu_data30), 
                       day120 = nsamples(otu_data120), 
                       day365 = nsamples(otu_data365))

# Find the most abundant taxa across time
top_phylum30 = top_taxa(phylum_data30, n = 50)
top_phylum120 = top_taxa(phylum_data120, n = 50)
top_phylum365 = top_taxa(phylum_data365, n = 50)

top_family30 = top_taxa(family_data30, n = 50)
top_family120 = top_taxa(family_data120, n = 50)
top_family365 = top_taxa(family_data365, n = 50)

common_phylum = Reduce(intersect, list(day30 = top_phylum30, 
                                       day120 = top_phylum120,
                                       day365 = top_phylum365))
common_phylum = common_phylum[!is.na(common_phylum)]
top_phylum = sort(common_phylum[1:10])

common_family = Reduce(intersect, list(day30 = top_family30, 
                                       day120 = top_family120,
                                       day365 = top_family365))
common_family = common_family[!is.na(common_family)]
top_family = sort(common_family[1:10])

# top_phylum = c("Acidobacteria", "Actinobacteria", "Bacteroidetes", "Cyanobacteria", "Firmicutes", 
#                "Fusobacteria", "Proteobacteria", "Tenericutes", "TM7", "Verrucomicrobia")
# top_family = c("Bacteroidaceae", "Bifidobacteriaceae", "Clostridiaceae", "Enterobacteriaceae", "Fusobacteriaceae",
#                "Lachnospiraceae", "Moraxellaceae", "Pasteurellaceae", "Ruminococcaceae", "Staphylococcaceae")

# Prevalence
# Phylum
qc_phylum = data.frame(matrix(NA, nrow = 10, ncol = 4))
colnames(qc_phylum) = c("phylum", "day30", "day120", "day365")

core_phylum30 = subset_taxa(phylum_data30, phylum %in% top_phylum)
core_phylum120 = subset_taxa(phylum_data120, phylum %in% top_phylum)
core_phylum365 = subset_taxa(phylum_data365, phylum %in% top_phylum)
qc_phylum$phylum = taxa_names(core_phylum30)
qc_phylum$day30 = apply(abundances(core_phylum30), 1, function(x) sum(x != 0)/nsamples(core_phylum30))
qc_phylum$day120 = apply(abundances(core_phylum120), 1, function(x) sum(x != 0)/nsamples(core_phylum120))
qc_phylum$day365 = apply(abundances(core_phylum365), 1, function(x) sum(x != 0)/nsamples(core_phylum365))

# Family
qc_family = data.frame(matrix(NA, nrow = 10, ncol = 4))
colnames(qc_family) = c("family", "day30", "day120", "day365")

core_family30 = subset_taxa(family_data30, family %in% top_family)
core_family120 = subset_taxa(family_data120, family %in% top_family)
core_family365 = subset_taxa(family_data365, family %in% top_family)
qc_family$family = taxa_names(core_family30)
qc_family$day30 = apply(abundances(core_family30), 1, function(x) sum(x != 0)/nsamples(core_family30))
qc_family$day120 = apply(abundances(core_family120), 1, function(x) sum(x != 0)/nsamples(core_family120))
qc_family$day365 = apply(abundances(core_family365), 1, function(x) sum(x != 0)/nsamples(core_family365))

# Outputs 
saveRDS(df_sample, file = "../data/nomic/qc/df_sample.rds")
saveRDS(df_sample, file = "../data/nomic/qc/qc_phylum.rds")
saveRDS(df_sample, file = "../data/nomic/qc/qc_family.rds")
```

# 2. Run SECOM {.tabset}

## Day 30

```{r, eval=FALSE}
# Phylum level
phyloseq1 = otu_data30
phyloseq2 = core_phylum30
pseqs = list(c(phyloseq1, phyloseq2))
pseudo = 0; zero_cut = 0.9; corr_cut = 0.5; lib_cut = 0
wins_quant = c(0.05, 0.95); method = "pearson"; soft = FALSE; thresh_len = 20
n_cv = 10; seed = 123; thresh_hard = 0.3; max_p = 0.005; n_cl = 5

linear_phy_30 = secom_linear(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                             wins_quant, method, soft, thresh_len, n_cv, 
                             seed, thresh_hard, max_p, n_cl)

R = 1000; max_p = 0.005
dist_phy_30 = secom_dist(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                         wins_quant, R, seed, max_p, n_cl)

# Family level
phyloseq1 = otu_data30
phyloseq2 = core_family30
pseqs = list(c(phyloseq1, phyloseq2))
pseudo = 0; zero_cut = 0.9; corr_cut = 0.5; lib_cut = 0
wins_quant = c(0.05, 0.95); method = "pearson"; soft = FALSE; thresh_len = 20
n_cv = 10; seed = 123; thresh_hard = 0.3; max_p = 0.005; n_cl = 5

linear_fam_30 = secom_linear(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                             wins_quant, method, soft, thresh_len, n_cv, 
                             seed, thresh_hard, max_p, n_cl)

max_p = 0.005
dist_fam_30 = secom_dist(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                         wins_quant, seed, max_p, n_cl)

# Get bias corrected abundances
# bias
s_30 = linear_phy_30$s_diff_hat

# Raw abundances
raw_phylum30 = abundances(phylum_data30)
raw_phylum30 = raw_phylum30[, names(s_30)]
raw_family30 = abundances(family_data30)
raw_family30 = raw_family30[, names(s_30)]

# Corrected abundances
o = log(raw_phylum30)
o[is.infinite(o)] = NA
d = nrow(o)
y_hat_phylum_30 = o - rowMeans(o, na.rm = TRUE) - t(replicate(d, s_30))

o = log(raw_family30)
o[is.infinite(o)] = NA
d = nrow(o)
y_hat_family_30 = o - rowMeans(o, na.rm = TRUE) - t(replicate(d, s_30))
```

## Day 120

```{r, eval=FALSE}
# Phylum level
phyloseq1 = otu_data120
phyloseq2 = core_phylum120
pseqs = list(c(phyloseq1, phyloseq2))
pseudo = 0; zero_cut = 0.9; corr_cut = 0.5; lib_cut = 0
wins_quant = c(0.05, 0.95); method = "pearson"; soft = FALSE; thresh_len = 20
n_cv = 10; seed = 123; thresh_hard = 0.3; max_p = 0.005; n_cl = 5

linear_phy_120 = secom_linear(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                             wins_quant, method, soft, thresh_len, n_cv, 
                             seed, thresh_hard, max_p, n_cl)

R = 1000; max_p = 0.005
dist_phy_120 = secom_dist(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                         wins_quant, R, seed, max_p, n_cl)

# Family level
phyloseq1 = otu_data120
phyloseq2 = core_family120
pseqs = list(c(phyloseq1, phyloseq2))
pseudo = 0; zero_cut = 0.9; corr_cut = 0.5; lib_cut = 0
wins_quant = c(0.05, 0.95); method = "pearson"; soft = FALSE; thresh_len = 20
n_cv = 10; seed = 123; thresh_hard = 0.3; max_p = 0.005; n_cl = 5

linear_fam_120 = secom_linear(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                             wins_quant, method, soft, thresh_len, n_cv, 
                             seed, thresh_hard, max_p, n_cl)

max_p = 0.005
dist_fam_120 = secom_dist(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                         wins_quant, seed, max_p, n_cl)

# Get bias corrected abundances
# bias
s_120 = linear_phy_120$s_diff_hat

# Raw abundances
raw_phylum120 = abundances(phylum_data120)
raw_phylum120 = raw_phylum120[, names(s_120)]
raw_family120 = abundances(family_data120)
raw_family120 = raw_family120[, names(s_120)]

# Corrected abundances
o = log(raw_phylum120)
o[is.infinite(o)] = NA
d = nrow(o)
y_hat_phylum_120 = o - rowMeans(o, na.rm = TRUE) - t(replicate(d, s_120))

o = log(raw_family120)
o[is.infinite(o)] = NA
d = nrow(o)
y_hat_family_120 = o - rowMeans(o, na.rm = TRUE) - t(replicate(d, s_120))
```

## Day 365

```{r, eval=FALSE}
# Phylum level
phyloseq1 = otu_data365
phyloseq2 = core_phylum365
pseqs = list(c(phyloseq1, phyloseq2))
pseudo = 0; zero_cut = 0.9; corr_cut = 0.5; lib_cut = 0
wins_quant = c(0.05, 0.95); method = "pearson"; soft = FALSE; thresh_len = 20
n_cv = 10; seed = 123; thresh_hard = 0.3; max_p = 0.005; n_cl = 5

linear_phy_365 = secom_linear(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                             wins_quant, method, soft, thresh_len, n_cv, 
                             seed, thresh_hard, max_p, n_cl)

R = 1000; max_p = 0.005
dist_phy_365 = secom_dist(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                         wins_quant, R, seed, max_p, n_cl)

# Family level
phyloseq1 = otu_data365
phyloseq2 = core_family365
pseqs = list(c(phyloseq1, phyloseq2))
pseudo = 0; zero_cut = 0.9; corr_cut = 0.5; lib_cut = 0
wins_quant = c(0.05, 0.95); method = "pearson"; soft = FALSE; thresh_len = 20
n_cv = 10; seed = 123; thresh_hard = 0.3; max_p = 0.005; n_cl = 5

linear_fam_365 = secom_linear(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                             wins_quant, method, soft, thresh_len, n_cv, 
                             seed, thresh_hard, max_p, n_cl)

max_p = 0.005
dist_fam_365 = secom_dist(pseqs, pseudo, zero_cut, corr_cut, lib_cut, 
                         wins_quant, seed, max_p, n_cl)

# Get bias corrected abundances
# bias
s_365 = linear_phy_365$s_diff_hat

# Raw abundances
raw_phylum365 = abundances(phylum_data365)
raw_phylum365 = raw_phylum365[, names(s_365)]
raw_family365 = abundances(family_data365)
raw_family365 = raw_family365[, names(s_365)]

# Corrected abundances
o = log(raw_phylum365)
o[is.infinite(o)] = NA
d = nrow(o)
y_hat_phylum_365 = o - rowMeans(o, na.rm = TRUE) - t(replicate(d, s_365))

o = log(raw_family365)
o[is.infinite(o)] = NA
d = nrow(o)
y_hat_family_365 = o - rowMeans(o, na.rm = TRUE) - t(replicate(d, s_365))
```

## Outputs

```{r, eval=FALSE}
saveRDS(linear_phy_30, file = "../data/nomic/phylum/linear_phy_30.rds")
saveRDS(dist_phy_30, file = "../data/nomic/phylum/dist_phy_30.rds")
saveRDS(linear_phy_120, file = "../data/nomic/phylum/linear_phy_120.rds")
saveRDS(dist_phy_120, file = "../data/nomic/phylum/dist_phy_120.rds")
saveRDS(linear_phy_365, file = "../data/nomic/phylum/linear_phy_365.rds")
saveRDS(dist_phy_365, file = "../data/nomic/phylum/dist_phy_365.rds")

saveRDS(linear_fam_30, file = "../data/nomic/familylinear_fam_30.rds")
saveRDS(dist_fam_30, file = "../data/nomic/familydist_fam_30.rds")
saveRDS(linear_fam_120, file = "../data/nomic/familylinear_fam_120.rds")
saveRDS(dist_fam_120, file = "../data/nomic/familydist_fam_120.rds")
saveRDS(linear_fam_365, file = "../data/nomic/familylinear_fam_365.rds")
saveRDS(dist_fam_365, file = "../data/nomic/familydist_fam_365.rds")

y_hat_phylum = list(day30 = y_hat_phylum_30,
                    day120 = y_hat_phylum_120,
                    day365 = y_hat_phylum_365)

y_hat_family = list(day30 = y_hat_family_30,
                    day120 = y_hat_family_120,
                    day365 = y_hat_family_365)

saveRDS(y_hat_phylum, file = "outputs/y_hat_phylum.rds")
saveRDS(y_hat_family, file = "outputs/y_hat_family.rds")
```

# 3. Run Sparcc

```{r, eval=FALSE}
set.seed(123)
R_hat_sparcc30 = sparcc(t(abundances(family_data30)))$Cor
dimnames(R_hat_sparcc30) = list(taxa_names(family_data30), taxa_names(family_data30))

R_hat_sparcc120 = sparcc(t(abundances(family_data120)))$Cor
dimnames(R_hat_sparcc120) = list(taxa_names(family_data120), taxa_names(family_data120))

R_hat_sparcc365 = sparcc(t(abundances(family_data365)))$Cor
dimnames(R_hat_sparcc365) = list(taxa_names(family_data365), taxa_names(family_data365))

sparcc_obj = list(day30 = R_hat_sparcc30, day120 = R_hat_sparcc120, day365 = R_hat_sparcc365)

saveRDS(sparcc_obj, file = "../data/nomic/sparcc/sparcc_obj.rds")
```

# 4. Data summary {.tabset}

## Sample size

```{r}
qc_sample = read_rds("../data/nomic/qc/df_sample.rds")
colnames(qc_sample) = c("Day 30", "Day 120", "Day 365")

datatable(qc_sample)
```

## Phylum prevalence

```{r}
qc_phylum = read_rds("../data/nomic/qc/qc_phylum.rds") %>%
  mutate_if(is.numeric, function(x) round(x, 2)) %>%
  arrange(phylum)
colnames(qc_phylum) = c("Phylum", "Day 30", "Day 120", "Day 365")

datatable(qc_phylum)
```

## Family prevalence

```{r}
qc_family = read_rds("../data/nomic/qc/qc_family.rds") %>%
  mutate_if(is.numeric, function(x) round(x, 2)) %>%
  arrange(family)
colnames(qc_family) = c("Family", "Day 30", "Day 120", "Day 365")

datatable(qc_family)
```

# 5. SECOM results: phylum level

## 5.1 Data import

```{r}
# Linear relationships
phylum_linear_time_30 = read_rds("../data/nomic/phylum/linear_phy_30.rds")
phylum_linear_time_120 = read_rds("../data/nomic/phylum/linear_phy_120.rds")
phylum_linear_time_365 = read_rds("../data/nomic/phylum/linear_phy_365.rds")
corr_phylum_linear_time_30 = phylum_linear_time_30$corr
corr_phylum_linear_time_120 = phylum_linear_time_120$corr
corr_phylum_linear_time_365 = phylum_linear_time_365$corr
p_phylum_linear_time_30 = phylum_linear_time_30$corr_p
p_phylum_linear_time_120 = phylum_linear_time_120$corr_p
p_phylum_linear_time_365 = phylum_linear_time_365$corr_p
cooccur_phylum_30 = phylum_linear_time_30$mat_cooccur
cooccur_phylum_120 = phylum_linear_time_120$mat_cooccur
cooccur_phylum_365 = phylum_linear_time_365$mat_cooccur

# Filter by p-values
alpha = 0.1
corr_phylum_linear_time_30[p_phylum_linear_time_30 >= alpha] = 0
corr_phylum_linear_time_120[p_phylum_linear_time_120 >= alpha] = 0
corr_phylum_linear_time_365[p_phylum_linear_time_365 >= alpha] = 0

# Filter by correlation coefficients
th = 0.1
corr_phylum_linear_time_30[abs(corr_phylum_linear_time_30) < th] = 0
corr_phylum_linear_time_120[abs(corr_phylum_linear_time_120) < th] = 0
corr_phylum_linear_time_365[abs(corr_phylum_linear_time_365) < th] = 0

# Filter by co-occurrence
overlap = 10
corr_phylum_linear_time_30[cooccur_phylum_30 < overlap] = 0
corr_phylum_linear_time_120[cooccur_phylum_120 < overlap] = 0
corr_phylum_linear_time_365[cooccur_phylum_365 < overlap] = 0

# Sort row and columns
corr_phylum_linear_time_30 = corr_phylum_linear_time_30[sort(rownames(corr_phylum_linear_time_30)),
                                                        sort(colnames(corr_phylum_linear_time_30))]
corr_phylum_linear_time_120 = corr_phylum_linear_time_120[sort(rownames(corr_phylum_linear_time_120)),
                                                          sort(colnames(corr_phylum_linear_time_120))]
corr_phylum_linear_time_365 = corr_phylum_linear_time_365[sort(rownames(corr_phylum_linear_time_365)),
                                                          sort(colnames(corr_phylum_linear_time_365))]

df_phylum_linear = data.frame(get_upper_tri(corr_phylum_linear_time_30)) %>%
  rownames_to_column("var1") %>%
  pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
  mutate(day = "Day 30") %>%
  bind_rows(
    data.frame(get_upper_tri(corr_phylum_linear_time_120)) %>%
    rownames_to_column("var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
    mutate(day = "Day 120")
  ) %>%
  bind_rows(
    data.frame(get_upper_tri(corr_phylum_linear_time_365)) %>%
    rownames_to_column("var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
    mutate(day = "Day 365")
  ) %>%
  filter(!is.na(value)) %>%
  mutate(value = round(value, 2),
         metric = "Pearson")

# Distance relationships
phylum_dist_time_30 = read_rds("../data/nomic/phylum/dist_phy_30.rds")
phylum_dist_time_120 = read_rds("../data/nomic/phylum/dist_phy_120.rds")
phylum_dist_time_365 = read_rds("../data/nomic/phylum/dist_phy_365.rds")
corr_phylum_dist_time_30 = phylum_dist_time_30$dcorr
corr_phylum_dist_time_120 = phylum_dist_time_120$dcorr
corr_phylum_dist_time_365 = phylum_dist_time_365$dcorr
p_phylum_dist_time_30 = phylum_dist_time_30$dcorr_p
p_phylum_dist_time_120 = phylum_dist_time_120$dcorr_p
p_phylum_dist_time_365 = phylum_dist_time_365$dcorr_p

# Filter by p-values
alpha = 0.1
corr_phylum_dist_time_30[p_phylum_dist_time_30 >= alpha] = 0
corr_phylum_dist_time_120[p_phylum_dist_time_120 >= alpha] = 0
corr_phylum_dist_time_365[p_phylum_dist_time_365 >= alpha] = 0

# Filter by correlation coefficients
th = 0.1
corr_phylum_dist_time_30[abs(corr_phylum_dist_time_30) < th] = 0
corr_phylum_dist_time_120[abs(corr_phylum_dist_time_120) < th] = 0
corr_phylum_dist_time_365[abs(corr_phylum_dist_time_365) < th] = 0

# Filter by co-occurrence
overlap = 10
corr_phylum_dist_time_30[cooccur_phylum_30 < overlap] = 0
corr_phylum_dist_time_120[cooccur_phylum_120 < overlap] = 0
corr_phylum_dist_time_365[cooccur_phylum_365 < overlap] = 0

# Sort row and columns
corr_phylum_dist_time_30 = corr_phylum_dist_time_30[sort(rownames(corr_phylum_dist_time_30)),
                                                    sort(colnames(corr_phylum_dist_time_30))]
corr_phylum_dist_time_120 = corr_phylum_dist_time_120[sort(rownames(corr_phylum_dist_time_120)),
                                                      sort(colnames(corr_phylum_dist_time_120))]
corr_phylum_dist_time_365 = corr_phylum_dist_time_365[sort(rownames(corr_phylum_dist_time_365)),
                                                      sort(colnames(corr_phylum_dist_time_365))]

df_phylum_dist = data.frame(get_upper_tri(corr_phylum_dist_time_30)) %>%
  rownames_to_column("var1") %>%
  pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
  mutate(day = "Day 30") %>%
  bind_rows(
    data.frame(get_upper_tri(corr_phylum_dist_time_120)) %>%
    rownames_to_column("var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
    mutate(day = "Day 120")
  ) %>%
  bind_rows(
    data.frame(get_upper_tri(corr_phylum_dist_time_365)) %>%
    rownames_to_column("var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
    mutate(day = "Day 365")
  ) %>%
  filter(!is.na(value)) %>%
  mutate(value = round(value, 2),
         metric = "Distance")

# Merge datasets
df_phylum = df_phylum_linear %>%
  bind_rows(
    df_phylum_dist
  )
phy_level = sort(union(df_phylum$var1, df_phylum$var2))
df_phylum$var1 = factor(df_phylum$var1, levels = phy_level)
df_phylum$var2 = factor(df_phylum$var2, levels = phy_level)
df_phylum$day = factor(df_phylum$day, levels = c("Day 30", "Day 120", "Day 365"))
df_phylum$metric = factor(df_phylum$metric, levels = c("Pearson", "Distance"))

# Bias corrected abundances
y_phylum_all = read_rds("../data/nomic/phylum/y_hat_phylum.rds")
y_phylum_time_30_all = exp(y_phylum_all$day30)
y_phylum_time_120_all = exp(y_phylum_all$day120)
y_phylum_time_365_all = exp(y_phylum_all$day365)

phylum_keep_time_30 = rownames(phylum_linear_time_30$y_hat)
phylum_keep_time_120 = rownames(phylum_linear_time_120$y_hat)
phylum_keep_time_365 = rownames(phylum_linear_time_365$y_hat)

ind30 = which(rownames(y_phylum_time_30_all) %in% phylum_keep_time_30)
y_phylum_time_30 = as.data.frame(y_phylum_time_30_all[ind30, ]) %>%
  bind_rows(as.data.frame(t(data.frame(Others = colSums(y_phylum_time_30_all[-ind30, ], na.rm = TRUE)))))

ind120 = which(rownames(y_phylum_time_120_all) %in% phylum_keep_time_120)
y_phylum_time_120 = as.data.frame(y_phylum_time_120_all[ind120, ]) %>%
  bind_rows(as.data.frame(t(data.frame(Others = colSums(y_phylum_time_120_all[-ind120, ], na.rm = TRUE)))))

ind365 = which(rownames(y_phylum_time_365_all) %in% phylum_keep_time_365)
y_phylum_time_365 = as.data.frame(y_phylum_time_365_all[ind365, ]) %>%
  bind_rows(as.data.frame(t(data.frame(Others = colSums(y_phylum_time_365_all[-ind365, ], na.rm = TRUE)))))
```

## 5.2 Visualization {.tabset}

### Correlation coefficients

```{r, fig.height=10, fig.width=13}
p_phylum_corr = df_phylum %>%
  ggplot(aes(var2, var1, fill = value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", na.value = "grey",
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name = NULL) +
  scale_x_discrete(drop = FALSE) +
  scale_y_discrete(drop = FALSE) +
  facet_grid(metric~day) +
  geom_text(aes(var2, var1, label = value), color = "black", size = 5) +
  labs(x = NULL, y = NULL, title = NULL) +
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),
        axis.text.y = element_text(size = 12),
        strip.text.x = element_text(size = 14),
        strip.text.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "top") +
  coord_fixed()

p_phylum_corr
```

### Mean abundances

```{r, fig.height=5, fig.width=10}
df_abn_phylum = y_phylum_time_30 %>%
  mutate_if(is.numeric, ~replace_na(., 0)) %>%
  mutate(ave = rowMeans(across(everything())),
         day = "30") %>%
  select(ave, day) %>%
  rownames_to_column("phylum") %>%
  bind_rows(
    y_phylum_time_120 %>%
      mutate_if(is.numeric, ~replace_na(., 0)) %>%
      mutate(ave = rowMeans(across(everything())),
             day = "120") %>%
      select(ave, day) %>%
      rownames_to_column("phylum")
  ) %>%
  bind_rows(
    y_phylum_time_365 %>%
      mutate_if(is.numeric, ~replace_na(., 0)) %>%
      mutate(ave = rowMeans(across(everything())),
             day = "365") %>%
      select(ave, day) %>%
      rownames_to_column("phylum")
  ) %>%
  filter(phylum != "Others")
df_abn_phylum$phylum = factor(df_abn_phylum$phylum, levels = phy_level)
df_abn_phylum$day = factor(df_abn_phylum$day, levels = c("30", "120", "365"))

p_phylum_abn = df_abn_phylum %>%
  ggplot(aes(x = day, y = ave, fill = phylum)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(x = "Day", y = "Abundance") +
  scale_fill_brewer(name = NULL, palette = "Set1", drop = FALSE) +
  scale_x_discrete(drop = FALSE) +
  facet_wrap(.~phylum, nrow = 2, scales = "free") +
  theme_bw()
p_phylum_abn
```

### Mean relative abundance

```{r}
df_rel_phylum = y_phylum_time_30 %>%
  mutate_if(is.numeric, ~replace_na(., 0)) %>%
  mutate(total = rowSums(across(everything())),
         day = "30",
         prop = total/sum(total)) %>%
  select(prop, day) %>%
  rownames_to_column("phylum") %>%
  bind_rows(
    y_phylum_time_120 %>%
      mutate_if(is.numeric, ~replace_na(., 0)) %>%
      mutate(total = rowSums(across(everything())),
             day = "120",
             prop = total/sum(total)) %>%
      select(prop, day) %>%
      rownames_to_column("phylum")
  ) %>%
  bind_rows(
    y_phylum_time_365 %>%
      mutate_if(is.numeric, ~replace_na(., 0)) %>%
      mutate(total = rowSums(across(everything())),
             day = "365",
             prop = total/sum(total)) %>%
      select(prop, day) %>%
      rownames_to_column("phylum")
  )

df_rel_phylum$phylum = factor(df_rel_phylum$phylum, levels = c(phy_level, "Others"))
df_rel_phylum$day = factor(df_rel_phylum$day, levels = c("30", "120", "365"))
p_phylum_rel = df_rel_phylum %>%
  ggplot(aes(x = day, y = prop, fill = phylum)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(x = "Day", y = "Relative Abundance") +
  scale_fill_brewer(name = NULL, palette = "Set1") +
  theme_bw()
p_phylum_rel

df_rel_phylum = df_rel_phylum %>%
  mutate(prop = round(prop, 2)) %>%
  pivot_wider(names_from = "day", values_from = "prop") %>%
  arrange(phylum)
datatable(df_rel_phylum)

# Combine abundance and relative abundance plots
p_phylum_abn = p_phylum_abn + 
  theme(legend.position = "none") +
  labs(x = NULL)
p_phylum_combine = ggarrange(p_phylum_abn, p_phylum_rel, nrow = 2,
                             heights = c(1, 1.5), labels = c("a", "b"))

ggsave(plot = p_phylum_combine, "../images/supp/supp_nomic_count_phylum.pdf", 
       height = 8, width = 10)   
ggsave(plot = p_phylum_combine, "../images/supp/supp_nomic_count_phylum.jpeg", 
       height = 8, width = 10, dpi = 300)
```

# 6. SECOM results: family level

## 6.1 Data import

```{r}
# Linear relationships
family_linear_time_30 = read_rds("../data/nomic/family/linear_fam_30.rds")
family_linear_time_120 = read_rds("../data/nomic/family/linear_fam_120.rds")
family_linear_time_365 = read_rds("../data/nomic/family/linear_fam_365.rds")
corr_family_linear_time_30 = family_linear_time_30$corr
corr_family_linear_time_120 = family_linear_time_120$corr
corr_family_linear_time_365 = family_linear_time_365$corr
p_family_linear_time_30 = family_linear_time_30$corr_p
p_family_linear_time_120 = family_linear_time_120$corr_p
p_family_linear_time_365 = family_linear_time_365$corr_p
cooccur_family_30 = family_linear_time_30$mat_cooccur
cooccur_family_120 = family_linear_time_120$mat_cooccur
cooccur_family_365 = family_linear_time_365$mat_cooccur

# Filter by p-values
alpha = 0.1
corr_family_linear_time_30[p_family_linear_time_30 >= alpha] = 0
corr_family_linear_time_120[p_family_linear_time_120 >= alpha] = 0
corr_family_linear_time_365[p_family_linear_time_365 >= alpha] = 0

# Filter by correlation coefficients
th = 0.1
corr_family_linear_time_30[abs(corr_family_linear_time_30) < th] = 0
corr_family_linear_time_120[abs(corr_family_linear_time_120) < th] = 0
corr_family_linear_time_365[abs(corr_family_linear_time_365) < th] = 0

# Filter by co-occurrence
overlap = 10
corr_family_linear_time_30[cooccur_family_30 < overlap] = 0
corr_family_linear_time_120[cooccur_family_120 < overlap] = 0
corr_family_linear_time_365[cooccur_family_365 < overlap] = 0

# Sort row and columns
corr_family_linear_time_30 = corr_family_linear_time_30[sort(rownames(corr_family_linear_time_30)),
                                                        sort(colnames(corr_family_linear_time_30))]
corr_family_linear_time_120 = corr_family_linear_time_120[sort(rownames(corr_family_linear_time_120)),
                                                          sort(colnames(corr_family_linear_time_120))]
corr_family_linear_time_365 = corr_family_linear_time_365[sort(rownames(corr_family_linear_time_365)),
                                                          sort(colnames(corr_family_linear_time_365))]

df_family_linear = data.frame(get_upper_tri(corr_family_linear_time_30)) %>%
  rownames_to_column("var1") %>%
  pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
  mutate(day = "Day 30") %>%
  bind_rows(
    data.frame(get_upper_tri(corr_family_linear_time_120)) %>%
    rownames_to_column("var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
    mutate(day = "Day 120")
  ) %>%
  bind_rows(
    data.frame(get_upper_tri(corr_family_linear_time_365)) %>%
    rownames_to_column("var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
    mutate(day = "Day 365")
  ) %>%
  filter(!is.na(value)) %>%
  mutate(value = round(value, 2),
         metric = "Pearson")

# Distance relationships
family_dist_time_30 = read_rds("../data/nomic/family/dist_fam_30.rds")
family_dist_time_120 = read_rds("../data/nomic/family/dist_fam_120.rds")
family_dist_time_365 = read_rds("../data/nomic/family/dist_fam_365.rds")
corr_family_dist_time_30 = family_dist_time_30$dcorr
corr_family_dist_time_120 = family_dist_time_120$dcorr
corr_family_dist_time_365 = family_dist_time_365$dcorr
p_family_dist_time_30 = family_dist_time_30$dcorr_p
p_family_dist_time_120 = family_dist_time_120$dcorr_p
p_family_dist_time_365 = family_dist_time_365$dcorr_p

# Filter by p-values
alpha = 0.1
corr_family_dist_time_30[p_family_dist_time_30 >= alpha] = 0
corr_family_dist_time_120[p_family_dist_time_120 >= alpha] = 0
corr_family_dist_time_365[p_family_dist_time_365 >= alpha] = 0

# Filter by correlation coefficients
th = 0.1
corr_family_dist_time_30[abs(corr_family_dist_time_30) < th] = 0
corr_family_dist_time_120[abs(corr_family_dist_time_120) < th] = 0
corr_family_dist_time_365[abs(corr_family_dist_time_365) < th] = 0

# Filter by co-occurrence
overlap = 10
corr_family_dist_time_30[cooccur_family_30 < overlap] = 0
corr_family_dist_time_120[cooccur_family_120 < overlap] = 0
corr_family_dist_time_365[cooccur_family_365 < overlap] = 0

# Sort row and columns
corr_family_dist_time_30 = corr_family_dist_time_30[sort(rownames(corr_family_dist_time_30)),
                                                    sort(colnames(corr_family_dist_time_30))]
corr_family_dist_time_120 = corr_family_dist_time_120[sort(rownames(corr_family_dist_time_120)),
                                                      sort(colnames(corr_family_dist_time_120))]
corr_family_dist_time_365 = corr_family_dist_time_365[sort(rownames(corr_family_dist_time_365)),
                                                      sort(colnames(corr_family_dist_time_365))]

df_family_dist = data.frame(get_upper_tri(corr_family_dist_time_30)) %>%
  rownames_to_column("var1") %>%
  pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
  mutate(day = "Day 30") %>%
  bind_rows(
    data.frame(get_upper_tri(corr_family_dist_time_120)) %>%
    rownames_to_column("var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
    mutate(day = "Day 120")
  ) %>%
  bind_rows(
    data.frame(get_upper_tri(corr_family_dist_time_365)) %>%
    rownames_to_column("var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
    mutate(day = "Day 365")
  ) %>%
  filter(!is.na(value)) %>%
  mutate(value = round(value, 2),
         metric = "Distance")

# Merge datasets
df_family = df_family_linear %>%
  bind_rows(
    df_family_dist
  )
fam_level = sort(union(df_family$var1, df_family$var2))
df_family$var1 = factor(df_family$var1, levels = fam_level)
df_family$var2 = factor(df_family$var2, levels = fam_level)
df_family$day = factor(df_family$day, levels = c("Day 30", "Day 120", "Day 365"))
df_family$metric = factor(df_family$metric, levels = c("Pearson", "Distance"))

# Bias corrected abundances
y_family_all = read_rds("../data/nomic/family/y_hat_family.rds")
y_family_time_30_all = exp(y_family_all$day30)
y_family_time_120_all = exp(y_family_all$day120)
y_family_time_365_all = exp(y_family_all$day365)

family_keep_time_30 = rownames(family_linear_time_30$y_hat)
family_keep_time_120 = rownames(family_linear_time_120$y_hat)
family_keep_time_365 = rownames(family_linear_time_365$y_hat)

ind30 = which(rownames(y_family_time_30_all) %in% family_keep_time_30)
y_family_time_30 = as.data.frame(y_family_time_30_all[ind30, ]) %>%
  bind_rows(as.data.frame(t(data.frame(Others = colSums(y_family_time_30_all[-ind30, ], na.rm = TRUE)))))

ind120 = which(rownames(y_family_time_120_all) %in% family_keep_time_120)
y_family_time_120 = as.data.frame(y_family_time_120_all[ind120, ]) %>%
  bind_rows(as.data.frame(t(data.frame(Others = colSums(y_family_time_120_all[-ind120, ], na.rm = TRUE)))))

ind365 = which(rownames(y_family_time_365_all) %in% family_keep_time_365)
y_family_time_365 = as.data.frame(y_family_time_365_all[ind365, ]) %>%
  bind_rows(as.data.frame(t(data.frame(Others = colSums(y_family_time_365_all[-ind365, ], na.rm = TRUE)))))
```

## 6.2 Visualization {.tabset}

### Correlation coefficients

```{r, fig.height=10, fig.width=13}
p_family_corr = df_family %>%
  ggplot(aes(var2, var1, fill = value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", na.value = "grey",
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name = NULL) +
  scale_x_discrete(drop = FALSE) +
  scale_y_discrete(drop = FALSE) +
  facet_grid(metric~day) +
  geom_text(aes(var2, var1, label = value), color = "black", size = 4) +
  labs(x = NULL, y = NULL, title = NULL) +
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1, 
                               title.position = "top", title.hjust = 0.5)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1, face = "italic"),
        axis.text.y = element_text(size = 12, face = "italic"),
        strip.text.x = element_text(size = 14),
        strip.text.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "top") +
  coord_fixed()

p_family_corr
ggsave(plot = p_family_corr, "../images/main/nomic_corr_family.pdf", height = 10, width = 13)   
ggsave(plot = p_family_corr, "../images/main/nomic_corr_family.jpeg", height = 10, width = 13, dpi = 300)
```

### Mean abundances

```{r, fig.height=5, fig.width=10}
df_abn_family = y_family_time_30 %>%
  mutate_if(is.numeric, ~replace_na(., 0)) %>%
  mutate(ave = rowMeans(across(everything())),
         day = "30") %>%
  select(ave, day) %>%
  rownames_to_column("family") %>%
  bind_rows(
    y_family_time_120 %>%
      mutate_if(is.numeric, ~replace_na(., 0)) %>%
      mutate(ave = rowMeans(across(everything())),
             day = "120") %>%
      select(ave, day) %>%
      rownames_to_column("family")
  ) %>%
  bind_rows(
    y_family_time_365 %>%
      mutate_if(is.numeric, ~replace_na(., 0)) %>%
      mutate(ave = rowMeans(across(everything())),
             day = "365") %>%
      select(ave, day) %>%
      rownames_to_column("family")
  ) %>%
  filter(family != "Others")
df_abn_family$family = factor(df_abn_family$family, levels = fam_level)
df_abn_family$day = factor(df_abn_family$day, levels = c("30", "120", "365"))

p_family_abn = df_abn_family %>%
  ggplot(aes(x = day, y = ave, fill = family)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(x = "Day", y = "Abundance") +
  scale_fill_brewer(name = NULL, palette = "Paired", drop = FALSE) +
  scale_x_discrete(drop = FALSE) +
  facet_wrap(.~family, nrow = 2, scales = "free") +
  theme_bw() +
  theme(legend.text = element_text(face = "italic"),
        strip.text = element_text(face = "italic"))
p_family_abn
```

### Mean relative abundance

```{r}
df_rel_family = y_family_time_30 %>%
  mutate_if(is.numeric, ~replace_na(., 0)) %>%
  mutate(total = rowSums(across(everything())),
         day = "30",
         prop = total/sum(total)) %>%
  select(prop, day) %>%
  rownames_to_column("family") %>%
  bind_rows(
    y_family_time_120 %>%
      mutate_if(is.numeric, ~replace_na(., 0)) %>%
      mutate(total = rowSums(across(everything())),
             day = "120",
             prop = total/sum(total)) %>%
      select(prop, day) %>%
      rownames_to_column("family")
  ) %>%
  bind_rows(
    y_family_time_365 %>%
      mutate_if(is.numeric, ~replace_na(., 0)) %>%
      mutate(total = rowSums(across(everything())),
             day = "365",
             prop = total/sum(total)) %>%
      select(prop, day) %>%
      rownames_to_column("family")
  )

df_rel_family$family = factor(df_rel_family$family, levels = c(fam_level, "Others"))
df_rel_family$day = factor(df_rel_family$day, levels = c("30", "120", "365"))
p_family_rel = df_rel_family %>%
  ggplot(aes(x = day, y = prop, fill = family)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(x = "Day", y = "Relative Abundance") +
  scale_fill_brewer(name = NULL, palette = "Paired") +
  theme_bw() +
  theme(legend.text = element_text(face = "italic"))
p_family_rel

df_rel_family = df_rel_family %>%
  mutate(prop = round(prop, 2)) %>%
  pivot_wider(names_from = "day", values_from = "prop") %>%
  arrange(family)
datatable(df_rel_family)

# Combine abundance and relative abundance plots
p_family_abn = p_family_abn + 
  theme(legend.position = "none") +
  labs(x = NULL)
p_family_combine = ggarrange(p_family_abn, p_family_rel, nrow = 2,
                             heights = c(1, 1.5), labels = c("a", "b"))

ggsave(plot = p_family_combine, "../images/main/nomic_count_family.pdf", 
       height = 8, width = 10)   
ggsave(plot = p_family_combine, "../images/main/nomic_count_family.jpeg", 
       height = 8, width = 10, dpi = 300)
```

### Scatter plots

```{r}
df_res = df_family %>%
  filter(value != 0) %>%
  pivot_wider(names_from = "metric", values_from = "value") %>%
  mutate(Pearson = replace_na(Pearson, 0),
         Distance = replace_na(Distance, 0)) %>%
  arrange(day, var1, var2)

y_family_list = list(day30 = family_linear_time_30$y_hat, 
                     day120 = family_linear_time_120$y_hat,
                     day365 = family_linear_time_365$y_hat)

scatter_family_time_30 = vector(mode = "list")
scatter_family_time_120 = vector(mode = "list")
scatter_family_time_365 = vector(mode = "list")
ind1 = 1; ind2 = 1; ind3 = 1
for (i in seq_along(df_res$var1)) {
  day = as.character(df_res$day[i])
  var1 = as.character(df_res$var1[i])
  var2 = as.character(df_res$var2[i])
  
  if (day == "Day 30") {
    df_y = y_family_list$day30
    df_fig = data.frame(y = df_y[var1, ], x = df_y[var2, ])
    
    fig_scatter = df_fig %>%
      ggplot(aes(x = x, y = y)) +
      labs(x = var2, y = var1) +
      geom_point() +
      geom_smooth(method = "loess", formula = "y ~ x", se = FALSE, color = "red") +
      geom_smooth(method = "lm", formula = "y ~ x", se = FALSE, color = "blue", linetype = "dashed") +
      theme_bw() +
      theme(axis.title.x = element_text(face = "italic"),
            axis.title.y = element_text(face = "italic"))
    scatter_family_time_30[[ind1]] = fig_scatter
    ind1 = ind1 + 1
  } else if (day == "Day 120") {
    df_y = y_family_list$day120
    df_fig = data.frame(y = df_y[var1, ], x = df_y[var2, ])
    
    fig_scatter = df_fig %>%
      ggplot(aes(x = x, y = y)) +
      labs(x = var2, y = var1) +
      geom_point() +
      geom_smooth(method = "loess", formula = "y ~ x", se = FALSE, color = "red") +
      geom_smooth(method = "lm", formula = "y ~ x", se = FALSE, color = "blue", linetype = "dashed") +
      theme_bw() +
      theme(axis.title.x = element_text(face = "italic"),
            axis.title.y = element_text(face = "italic"))
    scatter_family_time_120[[ind2]] = fig_scatter
    ind2 = ind2 + 1
  } else {
    df_y = y_family_list$day365
    df_fig = data.frame(y = df_y[var1, ], x = df_y[var2, ])
    
    fig_scatter = df_fig %>%
      ggplot(aes(x = x, y = y)) +
      labs(x = var2, y = var1) +
      geom_point() +
      geom_smooth(method = "loess", formula = "y ~ x", se = FALSE, color = "red") +
      geom_smooth(method = "lm", formula = "y ~ x", se = FALSE, color = "blue", linetype = "dashed") +
      theme_bw() +
      theme(axis.title.x = element_text(face = "italic"),
            axis.title.y = element_text(face = "italic"))
    scatter_family_time_365[[ind3]] = fig_scatter
    ind3 = ind3 + 1
  }
}

write.xlsx(df_res, "../outputs/nomic/secom_results.xlsx", overwrite = TRUE)

pair_scatter_time_30 = ggarrange(plotlist = scatter_family_time_30)
pair_scatter_time_120 = ggarrange(plotlist = scatter_family_time_120)
pair_scatter_time_365 = ggarrange(plotlist = scatter_family_time_365)

ggsave(plot = pair_scatter_time_30, "../images/supp/supp_nomic_scatter_family_30.pdf", 
       height = 3, width = 6)   
ggsave(plot = pair_scatter_time_30, "../images/supp/supp_nomic_scatter_family_30.jpeg", 
       height = 3, width = 6, dpi = 300)

ggsave(plot = pair_scatter_time_120, "../images/supp/supp_nomic_scatter_family_120.pdf", 
       height = 6, width = 6)   
ggsave(plot = pair_scatter_time_120, "../images/supp/supp_nomic_scatter_family_120.jpeg", 
       height = 6, width = 6, dpi = 300)

ggsave(plot = pair_scatter_time_365, "../images/supp/supp_nomic_scatter_family_365.pdf", 
       height = 8, width = 8)   
ggsave(plot = pair_scatter_time_365, "../images/supp/supp_nomic_scatter_family_365.jpeg", 
       height = 8, width = 8, dpi = 300)

# Example of nonlinear relationship
df_y = y_family_list$day365
df_fig = data.frame(y = df_y["Ruminococcaceae", ], x = df_y["Staphylococcaceae", ])
df_fig_complete = df_fig[complete.cases(df_fig), ]
df_ann = data.frame(x = -0.5, y = 5, 
                    pcor = round(cor.test(df_fig$x, df_fig$y, method = "pearson")$estimate, 2),
                    dcor = round(dcor(df_fig_complete$x, df_fig_complete$y), 2)) %>%
  mutate(label = paste0("pCor = ", pcor, "\ndCor = ", dcor))

fig_scatter = df_fig %>%
  ggplot(aes(x = x, y = y)) +
  labs(x = "Staphylococcaceae", y = "Ruminococcaceae") +
  geom_point() +
  geom_label(data = df_ann, aes(x = x, y = y, label = label),
             color = "orange", size = 4, fontface = "bold") +
  geom_smooth(method = "loess", formula = "y ~ x", se = FALSE, color = "red") +
  geom_smooth(method = "lm", formula = "y ~ x", se = FALSE, color = "blue", linetype = "dashed") +
  theme_bw() +
  theme(axis.title.x = element_text(face = "italic"),
        axis.title.y = element_text(face = "italic"))

```

## 6.3 Nonlinearity Examples {.tabset}

### Pasteurellaceae vs. Bacteroidaceae at day 365

```{r}
df_eg1 = data.frame(x = family_linear_time_365$y_hat["Pasteurellaceae", ],
                    y = family_linear_time_365$y_hat["Bacteroidaceae", ])
df_eg1 = df_eg1[complete.cases(df_eg1), ]

rlm1 = lmrob(y ~ x, data = df_eg1, control = lmrob.control(max.it = 100))
rlm2 = lmrob(y ~ poly(x, 2), data = df_eg1, control = lmrob.control(max.it = 100))
r_squared1 = round(summary(rlm1)$adj.r.squared, 2)
r_squared2 = round(summary(rlm2)$adj.r.squared, 2)

df_ann2 = data.frame(x = c(0.75, 0.75), y = c(1.75, 1.25)) %>%
  mutate(label = c(paste0("R^{2}~(linear) == ",  r_squared1),
                   paste0("R^{2}~(Poly) == ",  r_squared2)))

scatter_eg1 = df_eg1 %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(size = 4) +
  geom_text(data = df_ann2, aes(x = x, y = y, label = label), hjust = 0, 
             color = "orange", size = 4, fontface = "bold", parse = TRUE) +
  labs(x = "Enterobacteriaceae", y = "Ruminococcaceae") +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE, linetype = "dashed") + 
  theme_bw()
  
scatter_eg1
ggsave(plot = scatter_eg1, "../images/supp/supp_nonlinear_example1.pdf", height = 5, width = 6.25)
ggsave(plot = scatter_eg1, "../images/supp/supp_nonlinear_example1.jpeg", height = 5, width = 6.25, dpi = 300)
```

### Enterobacteriaceae vs. Ruminococcaceae at day 120

```{r}
df_eg2 = data.frame(x = family_linear_time_120$y_hat["Enterobacteriaceae", ],
                    y = family_linear_time_120$y_hat["Ruminococcaceae", ])
df_eg2 = df_eg2[complete.cases(df_eg2), ]

rlm1 = lmrob(y ~ x, data = df_eg2, control = lmrob.control(max.it = 100))
rlm2 = lmrob(y ~ poly(x, 4), data = df_eg2, control = lmrob.control(max.it = 100))
r_squared1 = round(summary(rlm1)$adj.r.squared, 2)
r_squared2 = round(summary(rlm2)$adj.r.squared, 2)

df_ann2 = data.frame(x = c(0.75, 0.75), y = c(1.75, 1.25)) %>%
  mutate(label = c(paste0("R^{2}~(linear) == ",  r_squared1),
                   paste0("R^{2}~(Poly) == ",  r_squared2)))

scatter_eg2 = df_eg2 %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(size = 4) +
  geom_text(data = df_ann2, aes(x = x, y = y, label = label), hjust = 0, 
             color = "orange", size = 4, fontface = "bold", parse = TRUE) +
  labs(x = "Enterobacteriaceae", y = "Ruminococcaceae") +
  geom_smooth(method = "lm", formula = y ~ poly(x, 4), se = FALSE, linetype = "dashed") + 
  theme_bw()
  
scatter_eg2
ggsave(plot = scatter_eg2, "../images/supp/supp_nonlinear_example2.pdf", height = 5, width = 6.25)
ggsave(plot = scatter_eg2, "../images/supp/supp_nonlinear_example2.jpeg", height = 5, width = 6.25, dpi = 300)
```

# 7. SparCC results

```{r, fig.height=6, fig.width=13}
top_phylum = c("Acidobacteria", "Actinobacteria", "Bacteroidetes", "Cyanobacteria", "Firmicutes", 
               "Fusobacteria", "Proteobacteria", "Tenericutes", "TM7", "Verrucomicrobia")
top_family = c("Bacteroidaceae", "Bifidobacteriaceae", "Clostridiaceae", "Enterobacteriaceae", "Fusobacteriaceae",
               "Lachnospiraceae", "Moraxellaceae", "Pasteurellaceae", "Ruminococcaceae", "Staphylococcaceae")

sparcc_obj = read_rds("../data/nomic/sparcc/sparcc_obj.rds")
R_hat_sparcc30 = sparcc_obj$day30
R_hat_sparcc120 = sparcc_obj$day120
R_hat_sparcc365 = sparcc_obj$day365

# Day 30
R_hat_sparcc30 = hard_thresh(R_hat_sparcc30, 0.3)
col_ind = match(sort(top_family), colnames(R_hat_sparcc30))
R_hat_sparcc30 = R_hat_sparcc30[col_ind[!is.na(col_ind)], col_ind[!is.na(col_ind)]]

# Day 120
R_hat_sparcc120 = hard_thresh(R_hat_sparcc120, 0.3)
col_ind = match(sort(top_family), colnames(R_hat_sparcc120))
R_hat_sparcc120 = R_hat_sparcc120[col_ind[!is.na(col_ind)], col_ind[!is.na(col_ind)]]

# Day 365
R_hat_sparcc365 = hard_thresh(R_hat_sparcc365, 0.3)
col_ind = match(sort(top_family), colnames(R_hat_sparcc365))
R_hat_sparcc365 = R_hat_sparcc365[col_ind[!is.na(col_ind)], col_ind[!is.na(col_ind)]]

df_sparcc = data.frame(get_upper_tri(R_hat_sparcc30)) %>%
  rownames_to_column("var1") %>%
  pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
  mutate(day = "Day 30") %>%
  bind_rows(
    data.frame(get_upper_tri(R_hat_sparcc120)) %>%
    rownames_to_column("var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
    mutate(day = "Day 120")
  ) %>%
  bind_rows(
    data.frame(get_upper_tri(R_hat_sparcc365)) %>%
    rownames_to_column("var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "value") %>%
    mutate(day = "Day 365")
  ) %>%
  filter(!is.na(value)) %>%
  mutate(value = round(value, 2))

fam_level = sort(union(df_sparcc$var1, df_sparcc$var2))
df_sparcc$var1 = factor(df_sparcc$var1, levels = fam_level)
df_sparcc$var2 = factor(df_sparcc$var2, levels = fam_level)
df_sparcc$day = factor(df_sparcc$day, levels = c("Day 30", "Day 120", "Day 365"))

p_sparcc_corr = df_sparcc %>%
  ggplot(aes(var2, var1, fill = value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", na.value = "grey",
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name = NULL) +
  scale_x_discrete(drop = FALSE) +
  scale_y_discrete(drop = FALSE) +
  facet_grid(.~day) +
  geom_text(aes(var2, var1, label = value), color = "black", size = 4) +
  labs(x = NULL, y = NULL, title = NULL) +
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1, 
                               title.position = "top", title.hjust = 0.5)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1, face = "italic"),
        axis.text.y = element_text(size = 12, face = "italic"),
        strip.text.x = element_text(size = 14),
        strip.text.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "top") +
  coord_fixed()

p_sparcc_corr
ggsave(plot = p_sparcc_corr, "../images/supp/supp_nomic_sparcc.pdf", height = 6, width = 13)
ggsave(plot = p_sparcc_corr, "../images/supp/supp_nomic_sparcc.jpeg", height = 6, width = 13, dpi = 300)
```

# Session information

```{r, message = FALSE, warning = FALSE, comment = NA}
sessionInfo()
```









